---
editor_options: 
  markdown: 
    wrap: 72
---

# Exercise : 02

## DATA TYPE - R OBJECT AND ATTRIBUTES DATATYPE VECTOR AND LIST

# 1. Basic Data Types in R

### Numeric

```{r}
num_var <- 10.5 
print(num_var)
class(num_var)

```

### Integer

```{r}
int_var <- 10L
print(int_var)
class(int_var)
```

### Character

```{r}
char_var <- "Hello R" 
print(char_var)
```

### Logical

```{r}
log_var <- TRUE 
print(log_var)
class(log_var)
```

### Complex

```{r}
comp_var <- 3 + 2i 
print(comp_var)
class(comp_var)
```

# 2. R Objects and Attributes

### Create a vector with attributes

```{r}
vec <- c(1, 2, 3, 4)
attr(vec, "description") <- "This is a numeric vector"
print(vec)
attributes(vec)
```

### Create a matrix with attributes

```{r}
mat <- matrix(1:9, nrow=3)
dimnames(mat) <- list(c("Row1", "Row2", "Row3"), c("Col1", "Col2", "Col3"))
print(mat)
attributes(mat)
```

# 3. Vectors in R

### Numeric vector

```{r}
num_vec <- c(1, 2, 3, 4, 5)
print(num_vec)
```

### Character vector

```{r}
char_vec <- c("apple", "banana", "cherry")
print(char_vec)
```

### Logical vector

```{r}
log_vec <- c(TRUE, FALSE, TRUE, FALSE)
print(log_vec)
```

### Operations on vectors

```{r}
vec1 <- c(1, 2, 3)
vec2 <- c(4, 5, 6)

sum_vec <- vec1 + vec2 # Element-wise addition
print(sum_vec)
```

# 4. Lists in R

### Creating a list with different data types

```{r}
my_list <- list(names = c("Dhayanidhi","Dhinesh"), age = c(22,20), marks = c(85, 90, 95))
print(my_list)
```

### Accessing elements in a list

```{r}
print(my_list$name) # Access by name
print(my_list[[3]]) # Access by index
```

### Modifying a list

```{r}
my_list$age <- 21
my_list$name <- "mahalakshmi"
print(my_list)
```

# Exercise : 03

## Data Types – Data Frame, Matrices, Factors, and Functions

### Data Frames

```{r}
df <- data.frame(Name = c("Alice", "Bob", "Charlie"), Age = c(25, 30, 35), Score = c(90, 85, 88))
print(df)
```

### Matrices

```{r}
a <- matrix(1:12, nrow = 3, ncol = 4)
b <- matrix(11:22, nrow = 3, ncol = 4)
print(a+b)
print(a-b)
print(a*b)
print(a/b)
```

### Factors

```{r}
gender <- factor(c("Male", "Female", "Male", "Female","Others"))
print(gender)
levels(gender)
```

### Functions

```{r}
square <- function(x) {
  return(x^2)
}
print(square(5))
```

```{r}
my_function <- function(x, y) {
  return(x + y)
}
print(my_function(3, 5))
```

# Exercise : 04

## EXPRESSION AND LOGICAL STATEMENT IN R

### Logical Expressions

```{r}
a <- 10
b <- 20
print(a > b)
print(a == b)
print(a != b)
print(a < b)

```

### Conditional Statements

```{r}
x <- 5
if (x > 0) {
  print("Positive number")
} else {
  print("Non-positive number")
}
```

### Looping with Logical Conditions

```{r}
for (i in 1:5) {
  if (i %% 2 == 0) {
    print(paste(i, "is even"))
  } else {
    print(paste(i, "is odd"))
  }
}
```

# Exercise : 05

## SUBSETTING OF LIST , MATRIX AND DATA FRAME

### Subsetting Lists

```{r}
new_list <- list(animal = "Tiger", count = 5, colors = c("Orange", "Black"))
print(new_list$animal) # By name
print(new_list[[2]]) # By index
print(new_list[1:2]) # Subsetting multiple elements
```

### Subsetting Matrices

```{r}
new_mat <- matrix(1:12, nrow=3, byrow=TRUE)
print(new_mat[1,])  # First row
print(new_mat[,2])  # Second column
print(new_mat[2,3]) # Specific element
```

### Subsetting Data Frames

```{r}
new_df <- data.frame(City = c("Chennai", "Mumbai"), Population = c(8000000, 20000000), Area = c(426, 603))
print(new_df)
print(new_df$City)    # Selecting a column
print(new_df[1,])     # Selecting a row
print(new_df[1:2, 2]) # Selecting specific rows and column
```

# Exercise : 06

## DATA FRAME FUNCTION ON INBUILT DATASET

### Working with Inbuilt Dataset

```{r}
data(iris)  # Loading iris dataset
print(head(iris))  # Display first six rows
```

### Summary Statistics

```{r}
print(summary(iris))  # Summary statistics of the dataset
```

### Selecting Specific Columns

```{r}
print(iris$Species)  # Selecting a specific column
print(iris[, c("Sepal.Length", "Sepal.Width")])  # Selecting multiple columns
```

### Filtering Data

```{r}
filtered_data <- subset(iris, Species == "setosa")
print(head(filtered_data))
```

### Aggregating Data

```{r}
aggr_data <- aggregate(Sepal.Length ~ Species, data = iris, FUN = mean)
print(aggr_data)
```

### Adding a New Column

```{r}
iris$Sepal.Area <- iris$Sepal.Length * iris$Sepal.Width
print(head(iris))
```

### Removing a Column

```{r}
iris <- subset(iris, select = -Sepal.Area)
print(head(iris))
```

# Exercise : 07

## DPLYR FUNCTION ON RETAIL DATASET

```{r}
library(dplyr)
library(ggplot2)
```

### Using the `diamonds` Dataset from ggplot2

```{r}
data("diamonds")
diamonds_sample <- diamonds %>% select(cut, color, clarity, price, carat)
print(head(diamonds_sample))
```

### Selecting Specific Columns

```{r}
selected_data <- select(diamonds_sample, cut, price)
print(head(selected_data))
```

### Filtering Data

```{r}
filtered_data <- filter(diamonds_sample, price > 5000)
print(head(filtered_data))
```

### Arranging Data

```{r}
sorted_data <- arrange(diamonds_sample, desc(price))
print(head(sorted_data))
```

### Mutating Data (Adding a New Column)

```{r}
diamonds_sample <- mutate(diamonds_sample, price_per_carat = price / carat)
print(head(diamonds_sample))
```

### Summarizing Data

```{r}
sales_summary <- diamonds_sample %>% summarise(Average_Price = mean(price))
print(sales_summary)
```

### Grouping and Summarizing Data

```{r}
grouped_summary <- diamonds_sample %>% group_by(cut) %>% summarise(Total_Price = sum(price))
print(grouped_summary)
```

# Exercise : 08

## DPLYR FUNCTION ON BANKING DATASET

### Loading Required Library

```{r}
library(dplyr)
```

### Creating a Sample Banking Dataset

```{r}
banking_data <- data.frame(
  Customer_ID = 1:6,
  Account_Type = c("Savings", "Checking", "Savings", "Loan", "Checking", "Loan"),
  Balance = c(5000, 2000, 7000, -1000, 1500, -500),
  Transactions = c(10, 25, 12, 5, 20, 8)
)
print(head(banking_data))

```

### Selecting Specific Columns

```{r}
selected_data <- select(banking_data, Customer_ID, Balance)
print(head(selected_data))
```

### Filtering Data

```{r}
filtered_data <- filter(banking_data, Balance > 1000)
print(head(filtered_data))
```

### Arranging Data

```{r}
sorted_data <- arrange(banking_data, desc(Balance))
print(head(sorted_data))
```

### Mutating Data (Adding a New Column)

```{r}
banking_data <- mutate(banking_data, Avg_Transaction = Balance / Transactions)
print(head(banking_data))
```

### Summarizing Data

```{r}
sales_summary <- banking_data %>% summarise(Average_Balance = mean(Balance))
print(sales_summary)
```

### Grouping and Summarizing Data

```{r}
grouped_summary <- banking_data %>% group_by(Account_Type) %>% summarise(Total_Balance = sum(Balance))
print(grouped_summary)
```

# Exercise : 09

## BASIC PLOTTING WITH R

```{r}

### Loading Required Library
library(ggplot2)

### Loading the mtcars Dataset
data("mtcars")
print(head(mtcars))

# 1. Histogram of MPG (Miles Per Gallon)
hist_plot <- ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(binwidth = 2, fill = "blue", color = "black") +
  labs(title = "Histogram of MPG", x = "Miles Per Gallon (mpg)", y = "Count")

# 2. Scatter Plot of Horsepower vs MPG
scatter_plot <- ggplot(mtcars, aes(x = hp, y = mpg)) +
  geom_point(color = "red") +
  labs(title = "Horsepower vs MPG", x = "Horsepower (hp)", y = "Miles Per Gallon (mpg)")

# 3. Bar Chart of Cylinders Count
bar_chart <- ggplot(mtcars, aes(x = factor(cyl))) +
  geom_bar(fill = "green") +
  labs(title = "Cylinder Count Distribution", x = "Number of Cylinders", y = "Count")

# 4. Box Plot of MPG by Number of Cylinders
box_plot <- ggplot(mtcars, aes(x = factor(cyl), y = mpg)) +
  geom_boxplot(fill = "purple") +
  labs(title = "MPG Distribution by Cylinders", x = "Number of Cylinders", y = "Miles Per Gallon (mpg)")

# 5. Line Plot of Weight vs MPG (Trend)
line_plot <- ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_line(color = "orange") +
  labs(title = "Weight vs MPG Trend", x = "Weight (wt)", y = "Miles Per Gallon (mpg)")

# Printing the Plots
print(hist_plot)
print(scatter_plot)
print(bar_chart)
print(box_plot)
print(line_plot)

```

# Exercise : 10

## GGPLOTS WITH R

```{r}

### Loading Required Libraries
library(ggplot2)
library(palmerpenguins)

### Loading the Penguins Dataset
data("penguins")
print(head(penguins))

# 1. Histogram of Flipper Length
hist_plot <- ggplot(penguins, aes(x = flipper_length_mm)) +
  geom_histogram(binwidth = 5, fill = "blue", color = "black") +
  labs(title = "Histogram of Flipper Length", x = "Flipper Length (mm)", y = "Count")

# 2. Scatter Plot of Bill Length vs Bill Depth
scatter_plot <- ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm, color = species)) +
  geom_point() +
  labs(title = "Bill Length vs Bill Depth", x = "Bill Length (mm)", y = "Bill Depth (mm)")

# 3. Bar Chart of Species Count
bar_chart <- ggplot(penguins, aes(x = species, fill = species)) +
  geom_bar() +
  labs(title = "Species Count", x = "Penguin Species", y = "Count")

# 4. Box Plot of Body Mass by Species
box_plot <- ggplot(penguins, aes(x = species, y = body_mass_g, fill = species)) +
  geom_boxplot() +
  labs(title = "Body Mass Distribution by Species", x = "Penguin Species", y = "Body Mass (g)")

# 5. Line Plot of Average Flipper Length by Species
line_plot <- ggplot(penguins, aes(x = species, y = flipper_length_mm, color = species, group = 1)) +
  stat_summary(fun = mean, geom = "point", size = 4) +  # Add points for mean values
  stat_summary(fun = mean, geom = "line", size = 1) +   # Connect points with a line
  labs(title = "Average Flipper Length by Species", x = "Penguin Species", y = "Flipper Length (mm)")

# Printing the Plots
print(hist_plot)
print(scatter_plot)
print(bar_chart)
print(box_plot)
print(line_plot)

```
